---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

```{r setoptions, echo=FALSE,warning = FALSE}
library("knitr")
opts_chunk$set(fig.path='figure/', warning = FALSE)
```

## Loading and preprocessing the data  
```{r}
DT1<-read.csv("activity.csv")
```

Creating directory for plots  
```{r}
dir.create("figure")
```

Removing rows with NA steps value.  
```{r}
DT2<-DT1[!is.na(DT1$steps),]
```

## Histogram of the total number of steps taken each day
```{r Plot1, fig.height=4,results='hide'}
stepsSums<-aggregate(DT2$steps,by=list(DT2$date),FUN=sum)
names(stepsSums)<-c("date","steps")
hist(stepsSums$steps, col="blue", xlab = "Steps", main="Total number of steps")
```

## Calculating mean total number of steps taken per day?
```{r, results='asis'}
library(plyr)
mm<-ddply(stepsSums,.(),summarize, mean=mean(steps), median=median(steps))
library(xtable)
mmtable<-xtable(mm[c("mean","median")])
print(mmtable, type = "html")
```

## The average daily activity pattern
```{r Plot2, fig.height=4,results='hide'}
intervalsMeans<-with(DT2,aggregate(steps,by=list(interval),FUN=mean))
names(intervalsMeans)<-c("interval","steps")
with(intervalsMeans,plot(x=interval,y=steps,type="l",xlab="Interval",
                         ,ylab="Steps", main="Daily activity"))
```

## Imputing missing values
1. Printing number of rows with NA value in 'steps' column
```{r}
sum(is.na(DT1$steps))
```

2. Creating list for replacing NA by means for inervals
```{r}
meansList<-intervalsMeans$steps
names(meansList)<-intervalsMeans$interval
```

3. Creating new table with filled NA
```{r}
DT3<-DT1
naIndices<-is.na(DT3)
DT3[naIndices,"steps"]<-meansList[as.character(DT3[naIndices,]$interval)]
```

4. Creating a hist and stats for new table...  
```{r Plot3, fig.height=4, results='hide'}
stepsSums3<-aggregate(DT3$steps,by=list(DT3$date),FUN=sum)
names(stepsSums3)<-c("date","steps")
hist(stepsSums3$steps, col="blue", xlab = "Steps", main="Total number of steps")
```
As is easy to see the count of days which total number of steps varies between 10000 and 15000 has increased.  
  
5. Creating the table of new mean and new median.
```{r, results='asis'}
mm<-ddply(stepsSums3,.(),summarize, mean=mean(steps), median=median(steps))
mmtable<-xtable(mm[c("mean","median")])
print(mmtable, type = "html")
```


## Creating a new factor variable indicating whether a given date is a weekday or weekend day.
```{r Plot4, fig.height=4,results='hide'}
library(chron)
library(ggplot2)
DT3$date<-as.Date(DT3$date)
DT3$daytype[is.weekend(DT3$date)]<-"weekend"
DT3$daytype[!is.weekend(DT3$date)]<-"weekday"
DT3$daytype<-as.factor(DT3$daytype)
qplot(interval, steps, data=DT3, main="Daily activity", facets=daytype~., 
      stat="summary",fun.y="mean", geom="line")
```